---
title: "Statistical Analyses for Van Meter et al 2016"
output:
#  pdf_document:
#    toc: yes
  html_document:
    toc: yes
---


```{r eval=FALSE, echo=FALSE}
#ctrl-shift k to knit r markdown
#https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
# TeX needed for markdown
#Windows: MiKTeX (Complete) - http://miktex.org/2.9/setup
#(NOTE: Be sure to download the Complete rather than Basic installation)

#Mac OS X: TexLive 2013 (Full) - http://tug.org/mactex/
#(NOTE: Download with Safari rather than Chrome _strongly_ recommended)

#Linux: Use system package manager
```

Introduction
========================================================
Data set, statistical analyses, graphics and inference concerning effect of organic carbon in soil on dermal bioconcentration in amphibians. Analytes include active ingredients for commonly used pesticides (atrazine, imidacloprid, fipronil, triadimefon, pendimethalin).

Distribution
========================================================
This page is public at:
http://rpubs.com/puruckertom/rvm2016

The git repo holding this code is available at:
https://github.com/puruckertom/VanMeteretal2016_ple_v_ols

Computational environment
========================================================

```{r eval=FALSE, echo=FALSE}
#library(installr) #windows only currently
#updateR()
```
Version and installed libraries.
```{r eval=TRUE, echo=TRUE}
R.Version()$version.string
Sys.info()[4]
#check to see if needed packages are installed
list.of.packages <- c("ggplot2", "dplyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#load relevant packages
library(MASS)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
```

```{r eval=TRUE, echo=FALSE}
if(Sys.info()[4]=="robins.computer.id (run 'Sys.info()[4]')"){
  frogsoildir <- "C:\\stp_drop\\Dropbox\\amphib_dermalexposure\\DATA\\PLEvsOLS2014\\"
}

if(Sys.info()[4]=="DC2626UTPURUCKE"){
  frogsoildir <- "C:\\stp_drop\\Dropbox\\amphib_ple_v_ols_vanmeteretal2015\\"
}

if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  frogsoildir <- "k:\\git\\VanMeteretal2016_ple_v_ols\\"
}

if(Sys.info()[4]=="stp-air.local"){
  frogsoildir <- path.expand("~/git/VanMeteretal2016_ple_v_ols/")
}

```


Experimental data
========================================================
Original file with analytical data: 
https://github.com/puruckertom/VanMeteretal2016_ple_v_ols/blob/master/RDATA.csv

```{r eval=TRUE, echo=TRUE}
##################
#the data sets
##################
#everything
file.exists(paste(frogsoildir,"RDATA.csv",sep=""))
frog.soil <- read.table(paste(frogsoildir,"RDATA.csv",sep=""), header = TRUE, sep = ",")
str(frog.soil)
```

Modifying data set to create factors for ANOVA, calculate BCFs and amphibian surface areas.
```{r eval=TRUE, echo=FALSE}
#class(frog.soil$Day)
frog.soil$Day <- as.factor(frog.soil$Day)
#frog.soil$Day
#class(frog.soil$Row)
frog.soil$Row <- as.factor(frog.soil$Row)
#frog.soil$Row
#class(frog.soil$Column)
#class(frog.soil$Pesticide)
#class(frog.soil$SoilType)
#class(frog.soil$Soil)
#class(frog.soil$BodyBurden)
#class(frog.soil$Weight)
#class(frog.soil$Total)
frog.soil$Total <- as.factor(frog.soil$Total)
#frog.soil$Total
#class(frog.soil$Formulation)
frog.soil$Formulation <- as.factor(frog.soil$Formulation)
#frog.soil$Formulation
#class(frog.soil$Parent)
frog.soil$Parent <- as.factor(frog.soil$Parent)
#frog.soil$Parent
```

```{r eval=TRUE, echo=TRUE}
frog.soil$bowlbcf <- frog.soil$BodyBurden/frog.soil$Soil
frog.soil$surface_area_total <- 1.131 * frog.soil$Weight^0.579
frog.soil$surface_area_footprint <- 0.425 * frog.soil$Weight^0.85
```

Structure of revised data set.
```{r eval=TRUE, echo=TRUE}
str(frog.soil)
```

```{r eval=FALSE, echo=FALSE}
#create a data set and do a merge
#koc data
#experimental
#log koc ple
#Imidacloprid 2.556
#Atrazine 2.303
#Triadimefon 3.025
#Fipronil 4.242
#Pendimethalin 6.425
#
#log koc ols
#Imidacloprid3.645
#Atrazine2.634
#Triadimefon3.010
#Fipronil 2.864
#Pendimethalin 1.733
#
#literature koc
#Atrazine  2.24
#Triadimefon  2.71
#Fipronil	2.916453949
#Pendimethalin	3.7
#Imidacloprid  2.394451681
```

Koc and organic carbon content data
========================================================
Add experimental and literature Koc data to data set.
```{r eval=TRUE, echo=TRUE}
unique(frog.soil$Pesticide)
unique(frog.soil$SoilType)
Pesticide <- c("ATZTOT","FipTOT","Imid","TNDTOT","Pendi","ATZTOT","FipTOT","Imid","TNDTOT","Pendi")
expKoc <- c(2.303,4.242,2.556,3.025,6.425,2.634,2.864,3.645,3.01,1.733)
litKoc <- c(2.24,2.92,2.39,2.71,3.7,2.24,2.92,2.39,2.71,3.7)
SoilType <- c("PLE","PLE","PLE","PLE","PLE","OLS","OLS","OLS","OLS","OLS")

df.merge <- data.frame(Pesticide, expKoc, litKoc, SoilType)
#View(merge(frog.soil, df.merge, all = TRUE))
frog.soil <- (merge(frog.soil, df.merge, all = TRUE))
```

Add organic carbon content data to dataframe.
```{r eval=TRUE, echo=TRUE}
#  Soil Type	%OM	%OC
#Plott Series	PLE	14.138	8.200
#Everbard	EC	5.448	3.160
#Orangeburg loamy-sand	OLS	3.100	1.798

SoilType <- c("PLE","OLS")
OM <- c(14.138, 3.1)
OC <- c(8.2, 1.798)
df.merge2 <- data.frame(SoilType, OM, OC)
frog.soil <- (merge(frog.soil, df.merge2, all = TRUE))
```

```{r eval=TRUE, echo=TRUE}
str(frog.soil)
```

Summary statistics
========================================================
Calculate mean and standard deviations for soil and amphibian tissure residue concentrations.

```{r eval=TRUE, echo=TRUE}
str(frog.soil)
head(frog.soil)
#View(frog.soil.total.ai)
#using dplyr
frog.soil.group <- group_by(frog.soil, Pesticide, SoilType, Formulation, Parent)
str(frog.soil.group)
frog.soil.group
frog.soil.means <- summarise(frog.soil.group, 
            count = n(),
            FrogMean = mean(BodyBurden),
            FrogSD = sd(BodyBurden),
            SoilMean = mean(Soil),
            SoilSD = sd(Soil)
  )
frog.soil.means
#View(frog.soil.means)

#Merge means and other statistics back into larger file.
frog.soil <- merge(frog.soil,frog.soil.means)

#treatment bcf
frog.soil$treatbcf <- frog.soil$BodyBurden/frog.soil$SoilMean
```


Setup of the complete main data set frog.soil.
```{r eval=TRUE, echo=TRUE}
dim(frog.soil)
str(frog.soil)
```

Subset data
========================================================
Eliminate data that will not be used for this analysis and manuscript. Keep only total pesticide results and only for active ingredients, metabolites and formulations dropped.

```{r eval=TRUE, echo=FALSE}
#total only but with formulations
frog.soil.total <- frog.soil[which(frog.soil$Total==1),]
#View(frog.soil.total)
dim(frog.soil.total)
colnames(frog.soil.total)
```

```{r eval=TRUE, echo=FALSE}
#Formulation is unbalanced so drop formulation, parent ais only
frog.soil.total.ai <- frog.soil.total[which(frog.soil.total$Formulation==0 & 
                                              frog.soil.total$Parent==1),]
#View(frog.soil.total.ai)
dim(frog.soil.total.ai)
colnames(frog.soil.total.ai)
```

This is the main table of summary results used in the results section for bcf (BodyBurden, treatbcf), for the soil table (SoilMean, SoilSD) and for the figure (treatbcf).
```{r eval=TRUE, echo=FALSE}
frog.soil.total.ai[,c(1,2,8,10,22:24)]
```

```{r eval=TRUE, echo=FALSE}
#drop imidacloprid from ai analysis
frog.soil.total.noimid <- frog.soil.total.ai[-which(frog.soil.total.ai$Pesticide=="Imid"),]
#View(frog.soil.total.noimid)
#dim(frog.soil.total.noimid)
#colnames
#dropping imidocloprid is unnecessary
```

Statistical analyses
========================================================
Analysis of total analyte active ingredients concentration data set. Previously, two alternative methods for calculating bioconcentration factors were used: dividing the amphibian tissue residue concentration of each bowl by the soil concentration within that bowl (bowlbcfs) and also dividing the amphibian tissue residue concentration of each bowl be the mean of all the bowls within the treatment (treatbcfs).

A paired comparison design is implemented to examine the impacts of soil type and surface area. The paired comparison design is essentially a randomized block design where the blocking variable (pesticides) has size 2 and is therefore treated as a nuisance variable. This controls for the large variation in treatment application rates and uptake across the different pesticides tested. The paired comparison design on the bowl bcf finds soil type significant, but not amphibian surface area. 

We evaluate 3 different paired comparison designs: with bowlbcfs, where each frog tissue concentration is divided by its soil concentration, mean bcf (treatbcf), where each frog tissue concentration is divided by the mean of the soil concentrations for that treatment, and by body.burden where soil concentration is used as an additional covariate instead of in the divisor.

```{r eval=TRUE, echo=FALSE}
##################
#analyses
##################
#randomized block design for bowlbcfs
bowlbcf.total.aov <- aov(bowlbcf ~ Pesticide + SoilType + Formulation + 
                           surface_area_total + expKoc, data = frog.soil.total)
summary(bowlbcf.total.aov)

#randomized block design for bowlbcfs
treatbcf.total.aov <- aov(treatbcf ~ Pesticide + SoilType + Formulation + 
                           surface_area_total + expKoc, data = frog.soil.total)
summary(treatbcf.total.aov)
```

Bowl bcf results and figure.
```{r eval=TRUE, echo=TRUE}
# is imidacloprid being factored in correctly for the aov
bowlbcf.total.ai.aov <- aov(bowlbcf ~ Pesticide + SoilType + surface_area_total + expKoc, 
                              data = frog.soil.total.ai)
summary(bowlbcf.total.ai.aov)
```

This figue based on bowlbcfs is not used in the manuscript.
```{r eval=TRUE, echo=TRUE}
qplot(Pesticide, bowlbcf, fill=factor(SoilType), data=frog.soil.total.ai, geom="boxplot", position="dodge")+theme_bw()
```

Treatment bcf results and figure. These analysis of variance results and the figure are used in the Van Meter at al. document. Context for interpretation - http://www.bodowinter.com/tutorial/bw_anova_general.pdf

Significance of soiltype: p = 0.00165, F= 10.999, df1 = 1, df2 = 53.
```{r eval=TRUE, echo=TRUE}
treatbcf.total.ai.aov <- aov(treatbcf ~ Pesticide + SoilType + expKoc, 
                              data = frog.soil.total.ai)
summary(treatbcf.total.ai.aov)
model.tables(treatbcf.total.ai.aov)
```

Generate figure as png for the manuscript. ggplot boxplots are different than regular R boxplots:
http://docs.ggplot2.org/0.9.3.1/geom_boxplot.html

The upper and lower "hinges" correspond to the first and third quartiles (the 25th and 75th percentiles). The upper whisker extends from the hinge to the highest value that is within 1.5 * IQR of the hinge, where IQR is the inter-quartile range, or distance between the first and third quartiles. The lower whisker extends from the hinge to the lowest value within 1.5 * IQR of the hinge. Data beyond the end of the whiskers are outliers and plotted as points (as specified by Tukey).

```{r eval=TRUE, echo=TRUE}
#figure 1 of Van Meter et al. manuscript
#pdf(paste(frogsoildir,"rvm2016_fig1.pdf", sep=""))
#levels(frog.soil.total.ai$Pesticide)
#levels(droplevels(frog.soil.total.ai$Pesticide))
#frog.soil.total.ai$Pesticide = factor(frog.soil.total.ai$Pesticide,levels(
#  frog.soil.total.ai$Pesticide)[c("Imid", "ATZTOT", "TNDTOT", "FipTOT", "Pendi")])
frog.soil.total.ai$Pesticide <- factor(frog.soil.total.ai$Pesticide, levels = c("Imid", "ATZTOT", "TNDTOT", "FipTOT", "Pendi"), ordered = TRUE)
#levels(frog.soil.total.ai$Pesticide) <- c("Imid", "ATZTOT", "TNDTOT", "FipTOT", "Pendi")
#png(paste(frogsoildir,"rvm2016_fig1.png", sep=""))
#tiff(paste(frogsoildir,"rvm2016_fig1.tiff", sep=""))

pesticides <- c("Imidacloprid", "Atrazine", "Triadimefon", "Fipronil", "Pendimethalin")

jpeg(paste(frogsoildir,"rvm2016_fig1.jpg", sep=""))
  pesticides <- c("Imidacloprid", "Atrazine", "Triadimefon", "Fipronil", "Pendimethalin")
  qplot(x=Pesticide, y=treatbcf, fill=factor(SoilType), xlab="", ylab="Bioconcentration Factor", 
        data=frog.soil.total.ai, geom="boxplot", position="dodge") + theme_bw() +  
        scale_x_discrete(breaks=c("Imid", "ATZTOT", "TNDTOT", "FipTOT", "Pendi"),  labels=pesticides)         + guides(fill=guide_legend(title="Soil Type")) + theme(legend.position="top")
dev.off()

```

For the html.
```{r eval=TRUE, echo=TRUE}
  #pesticides <- c("Atrazine", "Fipronil", "Imidacloprid", "Pendimethalin", "Triadimefon")
  #reordering by kow
  pesticides <- c("Imidacloprid", "Atrazine", "Triadimefon", "Fipronil", "Pendimethalin")
  qplot(x=Pesticide, y=treatbcf, fill=factor(SoilType), xlab="", ylab="Bioconcentration Factor", 
        data=frog.soil.total.ai, geom="boxplot", position="dodge") + 
        theme_bw() +  
        scale_x_discrete(breaks=c("Imid", "ATZTOT", "TNDTOT", "FipTOT", "Pendi"),  labels=pesticides) + 
        guides(fill=guide_legend(title="Soil Type")) + theme(legend.position="top")
```

Some additional stuff we did not use
========================================================
A figure.
```{r eval=TRUE, echo=TRUE}
boxplot(treatbcf ~ Pesticide + SoilType, data =frog.soil.total.ai)
```

We considered dropping imidacloprid for analytical reasons but it was unnecessary.
```{r eval=TRUE, echo=TRUE}
# more significant if imidacloprid is dropped (but not necessary)
bowlbcf.total.noimid.aov <- aov(bowlbcf ~ Pesticide + SoilType + surface_area_total, 
                                  data = frog.soil.total.noimid)
summary(bowlbcf.total.noimid.aov)
treatbcf.total.noimid.aov <- aov(treatbcf ~ Pesticide + SoilType + surface_area_total, 
                                  data = frog.soil.total.noimid)
summary(treatbcf.total.noimid.aov)
```

Tukey as anova post-hoc proxy.
```{r eval=TRUE, echo=TRUE}
TukeyHSD(treatbcf.total.ai.aov)
```

Bartletts test results.
```{r eval=TRUE, echo=TRUE}
#Bartlett test to test the null hypothesis of equal group variances
bartlett.test(treatbcf ~ Pesticide, data =frog.soil.total.ai)
#no sale! for pesticides
bartlett.test(treatbcf ~ SoilType, data = frog.soil.total.ai)
#also rejected for soil type, but not grouped by pesticide
```

Bartletts test results.
```{r eval=TRUE, echo=TRUE}
#the oneway.test( ) applies a Welch correction for nonhomogeneity
oneway.test(treatbcf ~ Pesticide + SoilType, data =frog.soil.total.ai)
```

Kruskal test results.
```{r eval=TRUE, echo=TRUE}
#nonparameteric kruskal test
kruskal.test(treatbcf ~ Pesticide, data =frog.soil.total.ai)
kruskal.test(treatbcf ~ SoilType, data =frog.soil.total.ai)
```

We can also consider the pairwise t-test. For this we need the means of the pesticide treatements by soil for the test. Doesn't make any sense.

```{r eval=FALSE, echo=FALSE}
#pairwise.ttest()
```